Local real space operators for translating Stokes Q/U to E/B maps on the sphere

### 2017-09-20 article/sections/discussion.tex
updates to the section qu2equbqu section in numerical miplementation ; added a set of new figures of the spectra derived from stokes maps for E and B derived from differencing the total from local Stokes Q U estimates corresponding to E and B

>>>>

\item{The difference between the spectra derived from the local convolution maps and the reference spectra follow a definite pattern and maybe it possible to model this difference and correct for it. For this purpose one will have to model the two point correlation function for the E and B fields in terms of the Stokes Q and U maps.}

\item{We have not addressed the E to B leakage issue in this work.}

### 2017-09-20 article/sections/numerical_implementation.tex
updates to the section qu2equbqu section in numerical miplementation ; added a set of new figures of the spectra derived from stokes maps for E and B derived from differencing the total from local Stokes Q U estimates corresponding to E and B

At first glance, the above remarks seem to hold true only for the E mode spectra, since the errors in the B mode spectra seem to be significantly larger as seen in the middle row of \fig{fig:eb-spectra_rad_cutoff}. We understand that the higher error in the B mode spectra is primarily a combined effect of the low amplitude of the B mode spectrum as compared to the E mode spectrum and limited numerical accuracy of our simple convolution algorithm. We ran a number of tests to confirm these as reasons for the observed discrepant behaviour. 
%
\begin{figure}[!t] 
\centering
%\vspace{-5cm}
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2eb/ee-spectrum-radial-cutoff.pdf}}
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2eb/bb-spectrum-radial-cutoff.pdf}}\\[-6ex]
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2eb/relative-percentage-err-ee-spectrum-radial-cutoff.pdf}}
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2eb/relative-percentage-err-bb-spectrum-radial-cutoff.pdf}}\\[-6ex]
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2eb/eqmask/relative-percentage-err-ee-spectrum-radial-cutoff.pdf}}
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2eb/eqmask/relative-percentage-err-bb-spectrum-radial-cutoff.pdf}} \\[-2ex]
\caption{\textit{Top:} The figures on left and right depict the E mode spectra and the B mode spectra respectively. \textit{Middle:} These figures depict the relative percentage difference between the spectra derived from full sky, locally evaluated E \& B  maps and the reference spectra. \textit{Bottom:} These figures depict the same results as shown in the middle row, except that the regions with $|b|>30^\circ$ of the maps have been masked for this analysis. 
The colors indicate the spectra derived from imposing different radial cutoff as indicated by the legends. The gray bands indicate the cosmic variance scatter around the fiducial spectra. The dashed horizontal black lines mark the 2\% error.}
\label{fig:eb-spectra_rad_cutoff}
\end{figure}
%
We exchanged the E and B mode spectra, \revisit{(taking care to set $C_{\ell}^{TE}=0$)} and noted that the trends observed in the error plots are reversed, clearly indicating a dependence on the amplitude of the spectrum. We noted that the polar caps were particularly error prone, as seen in \fig{fig:eb-maps-compare}. We excised these error prone regions, specifically masking portions of the maps with $|b|>30^{\circ}$ to find that the errors on the B mode spectra revert to having very similar structure as that of the E mode spectra, as seen in the bottom row of \fig{fig:eb-spectra_rad_cutoff}. Finally we modify the use of standard Healpix routines to mimic the local convolution E \& B construction (see \app{sec:convolution_err} for details). The results from this exercise also concur with our understanding of the discrepant behaviour for the B-mode spectrum.
%--------------------------------------------------------
%--------------------------------------------------------
\subsection{Separating Stokes Q \& U maps corresponding to E \& B modes of polarization}
Here 
%
\begin{figure}[!h] 
\centering
\subfigure[E Q-map: Healpix]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/e-qmap-healpix.pdf}}
\subfigure[E Q-map: Local convolution]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/e-qmap-2beta.pdf}}
\subfigure[E Q-map: Difference]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/e-qmap-diff.pdf}}
\subfigure[E U-map: Healpix]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/e-umap-healpix.pdf}}
\subfigure[E U-map: Local convolution]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/e-umap-2beta.pdf}}
\subfigure[E U-map: Difference]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/e-umap-diff.pdf}}

\subfigure[B Q-map: Healpix]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/b-qmap-healpix.pdf}}
\subfigure[B Q-map: Local convolution]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/b-qmap-2beta.pdf}}
\subfigure[B Q-map: Difference]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/b-qmap-diff.pdf}}
\subfigure[B U-map: Healpix]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/b-umap-healpix.pdf}}
\subfigure[B U-map: Local convolution]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/b-umap-2beta.pdf}}
\subfigure[B U-map: Difference]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/b-umap-diff.pdf}}
\caption{The top two rows depict the Stokes Q \& U maps  respectively corresponding to E mode of polarization while the bottom two rows depict  the Stokes Q \& U maps  respectively corresponding to the  B modes of polarization. \textit{Left:} Stokes Maps derived using Healpix. \textit{Middle}: Stokes maps derived using local convolution with $r_{\rm cutoff} = 2 \beta_o$. \textit{Right:} Difference between the maps derived using the two different methods.}
\label{fig:equ-bqu-maps-compare}
\end{figure}
%

%
\begin{figure}[!h] 
\centering
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/equ-spectra-radial-cutoff.pdf}}
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/bqu-spectra-radial-cutoff.pdf}}\\[-5ex]
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/relative-percentage-err-equ-e-spectrum-radial-cutoff.pdf}}
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/relative-percentage-err-bqu-b-spectrum-radial-cutoff.pdf}}
\caption{}
\label{fig:equ-bqu-spectra_rad_cutoff}
\end{figure}
%
%--------------------------------------------------------
%--------------------------------------------------------
\subsection{Multipole filtering using spatial convolutions}
%--------------------------------------------------------
%--------------------------------------------------------

>>>>

At first glance, the above remarks seem to hold true only for the E mode spectra, since the errors in the B mode spectra seem to be significantly larger as seen in the middle row of \fig{fig:eb-spectra_rad_cutoff}. We understand that the higher error in the B mode spectra is primarily a combined effect of the low amplitude of the B mode spectrum as compared to the E mode spectrum and limited numerical accuracy of our simple convolution algorithm. We ran a number of tests to confirm these as reasons for the observed discrepant behaviour. 
%
\begin{figure}[!t] 
\centering
%\vspace{-5cm}
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2eb/ee-spectrum-radial-cutoff.pdf}}
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2eb/bb-spectrum-radial-cutoff.pdf}}\\[-6ex]
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2eb/relative-percentage-err-ee-spectrum-radial-cutoff.pdf}}
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2eb/relative-percentage-err-bb-spectrum-radial-cutoff.pdf}}\\[-6ex]
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2eb/eqmask/relative-percentage-err-ee-spectrum-radial-cutoff.pdf}}
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2eb/eqmask/relative-percentage-err-bb-spectrum-radial-cutoff.pdf}} \\[-2ex]
\caption{\textit{Top:} The figures on left and right depict the E mode spectra and the B mode spectra respectively. \textit{Middle:} These figures depict the relative percentage difference between the spectra derived from full sky, locally evaluated E \& B  maps and the reference spectra. \textit{Bottom:} These figures depict the same results as shown in the middle row, except that the regions with $|b|>30^\circ$ of the maps have been masked for this analysis. 
The colors indicate the spectra derived from imposing different radial cutoff as indicated by the legends. The gray bands indicate the cosmic variance scatter around the fiducial spectra. The dashed horizontal black lines mark the 2\% error.}
\label{fig:eb-spectra_rad_cutoff}
\end{figure}
%
We exchanged the E and B mode spectra, \revisit{(taking care to set $C_{\ell}^{TE}=0$)} and noted that the trends observed in the error plots are reversed, clearly indicating a dependence on the amplitude of the spectrum. We noted that the polar caps were particularly error prone, as seen in \fig{fig:eb-maps-compare}. We excised these error prone regions, specifically masking portions of the maps with $|b|>30^{\circ}$ to find that the errors on the B mode spectra revert to having very similar structure as that of the E mode spectra, as seen in the bottom row of \fig{fig:eb-spectra_rad_cutoff}. Finally we modify the use of standard Healpix routines to mimic the local convolution E \& B construction (see \app{sec:convolution_err} for details). The results from this exercise also concur with our understanding of the discrepant behaviour for the B-mode spectrum.
%--------------------------------------------------------
%--------------------------------------------------------
\subsection{Separating Stokes Q \& U maps corresponding to E \& B modes of polarization}
In this section we present the results of evaluating the local convolution given in \eq{eq:op_qu2equbqu}, to decompose the total Stokes Q and U parameters into those correspond to the E modes and B modes of polarization respectively. We reiterate that this procedure uses real space operation on the total Stokes Q and U maps and does not go through the process of constructing the E and B modes as is usually the case. The resultant Stokes Q and U maps corresponding to the E mode are depicted in the first two rows while those corresponding to B modes are shown in the final two rows of \fig{fig:equ-bqu-maps-compare}. 
%
\begin{figure}[!t] 
\centering
\subfigure[E Q-map: Healpix]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/e-qmap-healpix.pdf}}
\subfigure[E Q-map: Local convolution]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/e-qmap-2beta.pdf}}
\subfigure[E Q-map: Difference]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/e-qmap-diff.pdf}}
\subfigure[E U-map: Healpix]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/e-umap-healpix.pdf}}
\subfigure[E U-map: Local convolution]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/e-umap-2beta.pdf}}
\subfigure[E U-map: Difference]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/e-umap-diff.pdf}}

\subfigure[B Q-map: Healpix]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/b-qmap-healpix.pdf}}
\subfigure[B Q-map: Local convolution]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/b-qmap-2beta.pdf}}
\subfigure[B Q-map: Difference]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/b-qmap-diff.pdf}}
\subfigure[B U-map: Healpix]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/b-umap-healpix.pdf}}
\subfigure[B U-map: Local convolution]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/b-umap-2beta.pdf}}
\subfigure[B U-map: Difference]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/b-umap-diff.pdf}}
\caption{The top two rows depict the Stokes Q \& U maps  respectively corresponding to E mode of polarization while the bottom two rows depict  the Stokes Q \& U maps  respectively corresponding to the  B modes of polarization. \textit{Left:} Stokes Maps derived using Healpix. \textit{Middle}: Stokes maps derived using local convolution with $r_{\rm cutoff} = 2 \beta_o$. \textit{Right:} Difference between the maps derived using the two different methods.}
\label{fig:equ-bqu-maps-compare}
\end{figure}
%
The local convolution maps depicted in this figure are computed using a radial cutoff of $r_{\rm cutoff }=2\beta_o$. Note that similar to observations of the previous section, the maps derived via local real space convolutions, suffer from having relatively higher error in the polar caps, while the differences in the  band around the equator are smaller and dominated by large scale modes.

Next we compute the E and B mode spectra from the set of Stokes maps constructed using local convolutions with different $r_{\rm cutoff}$. Note that for this exercise we use the standard Healpix routines for computing the spectra from the Stokes maps. The spectra derived from these local convolution maps and their difference with respect to the reference spectra are depicted in \fig{fig:equ-bqu-spectra_rad_cutoff}. The observations are similar to that made in the previous section: the difference with respect to the reference spectra reduce on increasing the radial cutoff and reach 2\% levels at most multipoles for the maximum radial $r_{\rm cutoff} =2 \beta_o$ used in this analysis.
%
\begin{figure}[!t] 
\centering
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/equ-spectra-radial-cutoff.pdf}}
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/bqu-spectra-radial-cutoff.pdf}}\\[-5ex]
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/relative-percentage-err-equ-e-spectrum-radial-cutoff.pdf}}
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/relative-percentage-err-bqu-b-spectrum-radial-cutoff.pdf}}
\caption{\textit{Top:} The thick lines in the figure on left depicts the E mode spectra derived from pure E Stokes Q and U maps derived using different radial cutoffs, while the thin lines depict the residual B mode power. The figure on the right depict the same except that the thick lines now depict the B mode spectra while the thin lines show the residual E mode power. \textit{Bottom:} The figure on the left and right depict the difference between the reference spectra and those derived from local convolution maps for different $r_{\rm cutoff}$ imposed on the convolution. In all the panels the color of the lines indicate the radial cutoff as indicated by the legend and the gray band marks the cosmic variance around the fiducial spectra.}
\label{fig:equ-bqu-spectra_rad_cutoff}
\end{figure}
%
In particular here we note that the B-mode residue in the Stokes Q \& U maps corresponding to E-modes is much smaller as compared to the E-mode residue in the Stokes Q \& U maps  computed for the B-modes of polarization. At all multipoles the B-mode residual power is below the lensing B-modes. The E-mode residual power at low multipoles is almost fully intact.

Motivated by these observations we also compute the Stokes Q \& U maps corresponding to B-modes of polarization by simply subtracting the Stokes Q \& U maps corresponding to E-mode from the total Stoker Q \& U maps. For completeness we also compute the Stokes Q \& U maps corresponding to E-mode in an analogous manner. We carry out a spectral study on these Stokes Q \& U maps corresponding to E \& B modes, computed using this alternate method. The results of the spectral analysis are summarized in \fig{fig:dbqu-spectra_rad_cutoff}.
%
\begin{figure}[!t] 
\centering
%\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/dequ-spectra-radial-cutoff.pdf}}
%\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/dbqu-spectra-radial-cutoff.pdf}} \\[-5ex]
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/relative-percentage-err-equ-de-spectrum-radial-cutoff.pdf}}
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/relative-percentage-err-bqu-db-spectrum-radial-cutoff.pdf}}
\caption{This figure depicts the same as the bottom panel of \fig{fig:equ-bqu-spectra_rad_cutoff}. The Stokes Q and U maps corresponding to E and B modes used in this analysis are constructed using the following relations: $\equ=\fqu-\fqu_{Local}^B$, $\bqu=\fqu-\fqu_{Local}^{E}$.}
\label{fig:dbqu-spectra_rad_cutoff}
\end{figure}
%
We observe that the the low multipole spectral behavior is not affected by this alternate construction of pure Stokes Q \& U maps. The high multipole behavior is significantly altered as can be seen by comparing \fig{fig:dbqu-spectra_rad_cutoff} with lower panels of \fig{fig:equ-bqu-spectra_rad_cutoff}. The relative percentage difference with respect to the reference spectra is seen to lie well within 2\% error for sufficiently high multipoles for all radial cutoff $r_{\rm cutoff}$.
%--------------------------------------------------------
%--------------------------------------------------------
\subsection{Multipole filtering using spatial convolutions}
%--------------------------------------------------------
%--------------------------------------------------------
